---
title: "South Carolina Department of Education data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Summary
This notebook loads, transforms and cleans data from South Carolina Department of Education’s annual report cards, head counts and additional data files for the years 2017-2021. Academic performance data based on state tests is not available for 2020. (Tests were canceled that year because of COVID-19.)

## Load libraries
```{r}
library(httr)
library(tidyverse)
library(jsonlite)
library(janitor)
library(parallel)
library(readxl)
library(rio)

```

## Define Functions
```{r}
## Function for standardizing column names
# Load master key of variable names
variable_names = read_excel("../data/handmade/variable_names.xlsx")

# Define function that maps variable names that should be replaced to their replacements
replace_names = function(df) {
  column_names = colnames(df)
  master_names_to_replace = c()
  master_new_name = c()
  for (i in column_names){
    
    names_to_replace = df[[i]]
    
    
    names_to_replace = names_to_replace[!is.na(names_to_replace)]
    new_names = rep(i, length(names_to_replace))
    
    master_names_to_replace = c(master_names_to_replace, names_to_replace)
    master_new_name = c(master_new_name, new_names)
    }
  
  master_key = setNames(nm = master_new_name, object = master_names_to_replace)
  
  return(master_key)
  
}

# Run function and store as list of keys
key = replace_names(variable_names)

```


## Load and clean data
### Create canonical district list using 2021 districts
```{r}

# Make a list of unneeded districts - Department of Juvenile Justice and Department of Corrections
drop_districts <- c("5208999", "5209999")

# Isolate district names in 2021 report cards
rc_district_list <- read_excel("../data/source/report-cards/rc_2021.xlsx", sheet = 3) %>% 
  clean_names() %>% 
  mutate(schoolid = as.character(schoolid)) %>% 
  rename(any_of(key)) %>% 
  # Filter only to districts
  filter(type == "D") %>%
  # Filter to exclude juvenile justice or correctional facilities
  filter(!(school_id %in% drop_districts)) %>%
  # Orangeburg County School District appears three times, all with the same info. Distinct df to remove two entries.
  distinct(school_id, .keep_all = TRUE) %>% 
  rename(district_id = school_id)

# There are 80 county-based districts, and five others: Charter Institute At Erskine, Governor's School For Agriculture At John De La Howe, SC Governor's School For The Arts And Humanities, SC Governor's School for Science and Mathematics, SC Public Charter School District

hc_district_list <- read_excel("../data/source/headcounts/2020_21_district_headcount_demographics.xlsx", col_names = c("district_id", "district", "total_actively_enrolled_students", "gender_female", "gender_male", "gender_missing", "black_or_african_american", "american_indian", "asian", "hispanic_or_latino", "hawaiian_or_other_pacific_islander", "two_or_more_races", "white", "race_missing", "pupils_in_poverty"), sheet = 1) %>%
  slice(8:94) %>%
  select(1:2) %>% 
  # Add trailing "999" to match ids in other datasets
  mutate(district_id = paste0(district_id, "999")) %>% 
  rename(district_alt = district)
  
# Create list with both versions of district name. Use an inner join to drop juvenile justice and corrections departments that are still in hc_district_list
district_list <- rc_district_list %>% 
  inner_join(hc_district_list) %>% 
  # Remove unneeded columns
  select(-headname, -year, -type)
  

# Create list with only the district names and ids for use in creation of school list
district_ids <- district_list %>% 
  select(district, district_alt, district_id)
  

```





### Create a canonical school list using 2021 schools
```{r}
# Make a list of unneeded entries - statewide and Department of Juvenile Justice
drop_schools <- c("9999999", "5208001", "5209001")

## How many schools are in each 2021 dataset?
# Load report card academic data for elementary and middle schools
rc_21_elem_mid <- read_excel("../data/source/report-cards/rc_2021.xlsx", sheet = 4) %>% 
  clean_names()

# Load report card data for high schools
rc_21_high <- read_excel("../data/source/report-cards/rc_2021.xlsx", sheet = 5) %>% 
  clean_names()

# Bind rows for report card academic data
rc_academic <- rc_21_elem_mid %>% 
  bind_rows(rc_21_high) %>%
  mutate(schoolid = as.character(schoolid)) %>% 
  rename(any_of(key)) %>% 
  filter(type != "D") %>% 
  filter(!(school_id %in% drop_schools)) %>% 
### This removes duplicate school ids, but we lose the distinction of some schools that have both elem and middle school data ("type" column)
  distinct(school_id, .keep_all = TRUE) %>% 
  mutate(len_str = nchar(school_id)) %>% 
  mutate(school_id = if_else(len_str == 6, paste0("0", school_id), school_id))

# Load report card main page
rc_main <- read_excel("../data/source/report-cards/rc_2021.xlsx", sheet = 3) %>% 
  clean_names() %>% 
  mutate(schoolid = as.character(schoolid)) %>% 
  rename(any_of(key)) %>% 
  # Filter to exclude districts, primary schools, the statewide entry
  filter(type != "D", type != "P") %>%
  # Filter to exclude two schools that are juvenile justice or correctional facilities
  filter(!(school_id %in% drop_schools)) %>% 
### This removes duplicate school ids, but we lose the distinction of some schools that have both elem and middle school data ("type" column) -- need to resolve
  distinct(school_id, .keep_all = TRUE)

# Load school headcount
school_hc_21 <- read_excel("../data/source/headcounts/2020_21_school_headcount_demographics.xlsx", col_names = c("school_id", "district", "school", "total_actively_enrolled_students", "gender_female", "gender_male", "gender_missing", "black_or_african_american", "american_indian", "asian", "hispanic_or_latino", "hawaiian_or_other_pacific_islander", "two_or_more_races", "white", "race_missing", "pupils_in_poverty"), sheet = 1) %>%
  slice(8:1227) %>%
  # Filter out schools that are part of the departments of juvenile justice or corrections
  filter(!grepl('Department', district)) %>% 
  mutate(year = "2021")

# The school list for additional info main is the same as report cards. We do not need to include this dataset in the joins for the school key.

  # Count number of rows in each dataset
  # nrow(rc_academic)
  # nrow(rc_main)
  # nrow(school_hc_21)
  
  # Check which schools are in rc_academic and not rc_main - none
  # anti <- rc_academic %>%
  #   anti_join(rc_main, by = c("year", "school_id"))
  
  # Check which schools are in rc_main and not rc_academic - Hcs Scholars Academy High School, which needs to be included in main school list
  # anti <- rc_main %>%
  #   anti_join(rc_academic, by = c("year", "school_id"))

## Join report card main list with report card academic list. Use left join to retain HCS Scholars Academy High School
rc_list <- rc_main %>%
  left_join(rc_academic, by = c("year", "school_id"))

# Join report card list with headcount list
school_list_rc_hc <- rc_list %>% 
  left_join(school_hc_21, by = "school_id") %>% 
  select(2:14,17) %>% 
  rename(district = district.x, school = school.x, type = type.x)


## Add charter schools
# Load and clean 2022 charter schools list. Sara Gregory created this file by exporting SCDE school directory data (https://ed.sc.gov/districts-schools/schools/school-directory/) and removing all schools except those that appeared on SCDE's website as a charter school as of April 15, 2022.
charter_schools_2022 <- read_csv("../data/handmade/charter_schools_04_15_2022.csv") %>% 
  clean_names %>% 
  # Create a column to identify these schools as charter schools
  mutate(type_special = "charter") %>% 
  # Change district and school code column types to characters
  mutate(districtcode = as.character(districtcode), schoolcode = as.character(schoolcode)) %>% 
  # Find length of district and school codes
  mutate(len_districtcode = nchar(districtcode), len_schoolcode = nchar(schoolcode)) %>% 
  # Add leading zeros to district and school ids
  mutate(districtcode = if_else(len_districtcode == 3, paste0("0", districtcode), districtcode)) %>% 
  # Add leading zeros to school ids
  mutate(schoolcode = if_else(len_schoolcode == 2, paste0("0", schoolcode), schoolcode)) %>% 
  mutate(schoolcode = if_else(len_schoolcode == 1, paste0("00", schoolcode), schoolcode)) %>% 
  # Create unique id by combining district and school ids. This will match the main school list
  mutate(school_id = paste0(districtcode, schoolcode))

# Create df with only the charter ids and the designation as a charter school
charter_ids <- charter_schools_2022 %>% 
  select(school_id, type_special)

# Join school list with charter school id list to add charter designation column.
school_list_with_charters <- school_list_rc_hc %>%
  left_join(charter_ids, by = "school_id")

    # Charter school checks
    # Check how many schools have been identified as charters in the school list
    # charters_in_school_list <- school_list %>%
    #   filter(type_special == "charter")
    
    # 16 schools for the 2022 charter list are not in the main school list. Identify those. 
    # missing_charters <- charter_schools_2022 %>%
    #   left_join(school_list, by = "school_id") %>%
    #   filter(is.na(school))


## Add district ids to school list
school_list <- district_ids %>% 
  left_join(school_list_with_charters, by = "district")

## Export school list for geocoding (in export chunk below)

```

```{r}
## Load geocoded school list
school_list_geocoded = read_csv("../data/processed/school_list_geocoded.csv")

```




### Load and clean school performance data
#### Academic data for elementary and middle schools
```{r}
#2017
## Load sheet with academic variables for elementary schools. Note: elem science scores for 2017 were not available at a school level.
acad_elem_17 <- read_excel("../data/source/report-cards/rc_2017_elem.xlsx", sheet = 2) %>% 
  clean_names() %>%
  mutate(e_pct_me = sc17_eng4e_xp + sc17_eng3r_ep, m_pct_me = sc17_mat4e_xp + sc17_mat3r_ep,
         year = "2017", id = as.character(id)) %>% 
  select("year", 2:5, "e_pct_me", "sc17_eng1n_sp", "m_pct_me", "sc17_mat1n_sp") %>% 
  rename(any_of(key)) 

########################## !!!!! Resolve "p17_sci1nm" issues
## Load sheet with academic variables for middle schools
acad_mid_17 <- read_excel("../data/source/report-cards/rc_2017_middle.xlsx", sheet = 2) %>% 
  clean_names() %>%
  mutate(e_pct_me = sc17_eng4e_xp + sc17_eng3r_ep, m_pct_me = sc17_mat4e_xp + sc17_mat3r_ep, sc_pct_me = p17_sci5exe + p17_sci7exc, year = "2017", id = as.character(id)) %>%
  select("year", 2:5, "e_pct_me", "sc17_eng1n_sp", "m_pct_me", "sc17_mat1n_sp", "sc_pct_me", "p17_sci1nm") %>% 
  rename(any_of(key))

# 2018
## Load sheet with academic variables for elementary and middle schools
acad_elemmid_18 <- read_excel("../data/source/report-cards/rc_2018.xlsx", sheet = 4) %>% 
  clean_names() %>% 
  select(1:5,"e_pct_me", "e_pct_in_need_of_support","m_pct_me", "m_pct_in_need_of_support", "sc_pct_me", "sc_pct_in_need_of_support") %>% 
  rename(any_of(key))


#2019
## Load sheet with academic variables for elementary and middle schools
acad_elemmid_19 <- read_excel("../data/source/report-cards/rc_2019.xlsx", sheet = 4) %>% 
  clean_names() %>% 
  select(1:5,"e_pct_me", "e_pct_in_need_of_support","m_pct_me", "m_pct_in_need_of_support", "sc_pct_me", "sc_pct_in_need_of_support") %>% 
   rename(any_of(key))

#2020
## No state assessments were given in spring 2020, so no metrics for English, math or science performance at elementary or middle schools

#2021
acad_elemmid_21 <- read_excel("../data/source/report-cards/rc_2021.xlsx", sheet = 4) %>% 
  clean_names() %>% 
  select(1:5,"e_pct_me", "e_pct_in_need_of_support","m_pct_me", "m_pct_in_need_of_support", "sc_pct_me", "sc_pct_in_need_of_support") %>% 
  rename(any_of(key)) %>% 
  # Convert school id and year to character. Convert performance metrics to numeric. In the latter change, schools with suppressed data are coerced from "*" to "NA".
  mutate(school_id = as.character(school_id), year = as.character(year), e_pct_me = as.numeric(e_pct_me), e_pct_in_need_of_support = as.numeric(e_pct_in_need_of_support), m_pct_me = as.numeric(m_pct_me), m_pct_in_need_of_support = as.numeric(m_pct_in_need_of_support), sc_pct_me = as.numeric(sc_pct_me), sc_pct_in_need_of_support = as.numeric(sc_pct_in_need_of_support))


# Bind elementary and middle schools across years
acad_em_bind = acad_elem_17 %>% 
  bind_rows(acad_mid_17) %>% 
  bind_rows(acad_elemmid_18) %>% 
  bind_rows(acad_elemmid_19) %>% 
  bind_rows(acad_elemmid_21) %>%
  mutate(len_str = nchar(school_id)) %>% 
  mutate(school_id = if_else(len_str == 6, paste0("0", school_id), school_id)) %>% 
  select(-len_str)

```


### Academic data for high schools
```{r}
# Define list of academic variables
acad_high_vars <- c("report_card_year", "district_nm", "school_nm", "schoolid", "schooltypecd", "e_pct_abc", "e_pct_ab", "e_pct_cd", "e_pct_f", "m_pct_abc", "m_pct_ab", "m_pct_cd", "m_pct_f", "sc_pct_abc", "sc_pct_ab", "sc_pct_cd", "sc_pct_f")


# 2017 - No academic grades for 2017. Metric changed between 2017 and 2018, so we cannot use comparable data from 2017.

acad_high_18 <- read_excel("../data/source/report-cards/rc_2018.xlsx", sheet = 5) %>% 
  clean_names() %>% 
  # English grades
  mutate(e_pct_ab = e_pct_a + e_pct_b, e_pct_cd = e_pct_c + e_pct_d) %>% 
  # Algebra grades
  mutate(sc_pct_ab= sc_pct_a + sc_pct_b, sc_pct_cd= sc_pct_c + m_pct_d) %>%
  # Biology grades
  mutate(m_pct_ab = m_pct_a + m_pct_b, m_pct_cd = m_pct_c + m_pct_d) %>%
  select(any_of(acad_high_vars))

  # Note: There are about 17 high schools or districts with 0 students scoring A or B in English ... is that accurate?

acad_high_19 <- read_excel("../data/source/report-cards/rc_2019.xlsx", sheet = 5) %>% 
  clean_names() %>% 
  # English grades
  mutate(e_pct_ab = e_pct_a + e_pct_b, e_pct_cd = e_pct_c + e_pct_d) %>% 
  # Algebra grades
  mutate(sc_pct_ab= sc_pct_a + sc_pct_b, sc_pct_cd= sc_pct_c + m_pct_d) %>%
  # Biology grades
  mutate(m_pct_ab = m_pct_a + m_pct_b, m_pct_cd = m_pct_c + m_pct_d) %>%
  select(any_of(acad_high_vars))

# No academic grades for 2020.

acad_high_21 <- read_excel("../data/source/report-cards/rc_2019.xlsx", sheet = 5) %>% 
  clean_names() %>% 
  # English grades
  mutate(e_pct_ab = e_pct_a + e_pct_b, e_pct_cd = e_pct_c + e_pct_d) %>% 
  # Algebra grades
  mutate(sc_pct_ab= sc_pct_a + sc_pct_b, sc_pct_cd= sc_pct_c + m_pct_d) %>%
  # Biology grades
  mutate(m_pct_ab = m_pct_a + m_pct_b, m_pct_cd = m_pct_c + m_pct_d) %>%
  select(any_of(acad_high_vars))


# Bind academic variables across years
acad_high_bind =  bind_rows(acad_high_18) %>% 
  bind_rows(acad_high_19) %>% 
  bind_rows(acad_high_21) %>%
  rename(any_of(key))
  # School ids are all 7 digits

```


### ACT scores
```{r}
# Define list of ACT variables
act_vars <- c("report_card_year", "district_nm", "school_nm", "schoolid", "schooltypecd", "act_avg_english_score", "act_avg_reading_score", "act_avg_math_score", "act_avg_science_score", "act_avg_writing_score", "act_avg_composite_score")


# 2017
# Create district df for 2017 that multiple climate variables will pull from
rc_dist_17 <- read_excel("../data/source/report-cards/rc_2017_district.xlsx", sheet = 2) %>% 
  clean_names() %>% 
  mutate(report_card_year = "2017", id = as.character(id)) %>% 
  rename(dropout_rate_curr_yr = "c16", act_avg_english_score = "a17_d_eng_avg", act_avg_reading_score = "a17_d_rea_avg", act_avg_math_score = "a17_d_mat_avg", act_avg_science_score = "a17_d_sci_avg", act_avg_writing_score = "a17_d_wrt_avg", act_avg_composite_score = "a17_d_com_avg", pct_act = "a17_d_comp20", district_nm = "distname", school_nm = "locname", schoolid = "id", schooltypecd = "type")

# Create high school df for 2017 that multiple climate variables will pull from
rc_high_17 <- read_excel("../data/source/report-cards/rc_2017_high.xlsx", sheet = 2) %>% 
  clean_names() %>% 
  mutate(report_card_year = "2017", id = as.character(id)) %>% 
  rename(dropout_rate_curr_yr = "c16", act_avg_english_score = "a17_eng_avg", act_avg_reading_score = "a17_rea_avg", act_avg_math_score = "a17_mat_avg", act_avg_science_score = "a17_sci_avg", act_avg_writing_score = "a17_wrt_avg", act_avg_composite_score = "a17_com_avg", pct_act = "a17_comp20", district_nm = "distname", school_nm = "locname", schoolid = "id", schooltypecd = "type")


act_dist_17 <- rc_dist_17 %>% 
  select(any_of(act_vars))

act_high_17 <- rc_high_17 %>% 
  select(any_of(act_vars))


# 2018
act_18 <- read_excel("../data/source/additional-info/ai-2018.xlsx", sheet = 5) %>% 
  clean_names() %>% 
  select(any_of(act_vars))

# 2019
act_19 <- read_excel("../data/source/additional-info/ai-2019.xlsx", sheet = 5) %>% 
  clean_names() %>% 
  select(any_of(act_vars))

# 2020
act_20 <- read_excel("../data/source/additional-info/ai-2020.xlsx", sheet = 6) %>% 
  clean_names() %>%
  mutate(report_card_year = as.character(report_card_year), schoolid = as.character(schoolid), act_avg_english_score = as.numeric(act_avg_english_score), act_avg_reading_score = as.numeric(act_avg_reading_score), act_avg_math_score = as.numeric(act_avg_math_score), act_avg_science_score = as.numeric(act_avg_science_score), act_avg_writing_score = as.numeric(act_avg_writing_score), act_avg_composite_score = as.numeric(act_avg_composite_score)) %>% 
  select(any_of(act_vars))

# 2021
act_21 <- read_excel("../data/source/additional-info/ai-2021.xlsx", sheet = 6) %>% 
  clean_names() %>% 
  mutate(report_card_year = as.character(report_card_year), schoolid = as.character(schoolid), act_avg_english_score = as.numeric(act_avg_english_score), act_avg_reading_score = as.numeric(act_avg_reading_score), act_avg_math_score = as.numeric(act_avg_math_score), act_avg_science_score = as.numeric(act_avg_science_score), act_avg_writing_score = as.numeric(act_avg_writing_score), act_avg_composite_score = as.numeric(act_avg_composite_score)) %>% 
  select(any_of(act_vars))


# Bind years
act_bind =  bind_rows(act_dist_17) %>% 
  bind_rows(act_high_17) %>% 
  bind_rows(act_18) %>% 
  bind_rows(act_19) %>%
  bind_rows(act_20) %>% 
  bind_rows(act_21) %>%
  rename(any_of(key))

```


### College and career readiness
```{r}
# Define list of college and career readiness variables
career_vars <- c("report_card_year", "district_nm", "school_nm", "schoolid", "schooltypecd", "pct_college", "pct_career", "pct_act", "pct_cte", "pct_wbl", "pct_car_rdy")

# 2017
career_dist_17 <- rc_dist_17 %>% 
  select(any_of(career_vars))
  # Only one variable available

career_high_17 <- rc_high_17 %>% 
  select(any_of(career_vars))
  # Only one variable available


# 2018
career_18 <- read_excel("../data/source/report-cards/rc_2018.xlsx", sheet = 6) %>% 
  clean_names() %>% 
  select(1:5, "pct_college", "pct_career", "pct_act", "pct_cte", "pct_wbl", "pct_car_rdy")

# 2019
career_19 <- read_excel("../data/source/report-cards/rc_2019.xlsx", sheet = 7) %>% 
  clean_names() %>% 
  select(1:5, "pct_college", "pct_career", "pct_act", "pct_cte", "pct_wbl", "pct_car_rdy")

# 2020
career_20 <- read_excel("../data/source/report-cards/rc_2020.xlsx", sheet = 5) %>% 
  clean_names() %>% 
  select(1:5, "pct_career", "pct_cte", "pct_wbl", "pct_car_rdy")
  # "% diploma earners that were college ready" and "% diploma earners who met the college ready threshold for ACT" are not included in 2020 report card

# 2021
career_21 <- read_excel("../data/source/report-cards/rc_2021.xlsx", sheet = 8) %>% 
  clean_names() %>%
  mutate(report_card_year = as.character(report_card_year), schoolid = as.character(schoolid), pct_college = as.numeric(pct_college), pct_career = as.numeric(pct_career), pct_act = as.numeric(pct_act), pct_cte = as.numeric(pct_cte), pct_wbl = as.numeric(pct_wbl), pct_car_rdy = as.numeric(pct_car_rdy)) %>%
  select(1:5, "pct_college", "pct_career", "pct_act", "pct_cte", "pct_wbl", "pct_car_rdy")


# Bind college and career readiness variables across years
career_bind =  bind_rows(career_dist_17) %>% 
  bind_rows(career_high_17) %>% 
  bind_rows(career_18) %>%
  bind_rows(career_19) %>% 
  bind_rows(career_21) %>%
  rename(any_of(key)) %>% 
  # Add leading zero to school ids with only 6 digits
  mutate(len_str = nchar(school_id)) %>% 
  mutate(school_id = if_else(len_str == 6, paste0("0", school_id), school_id)) %>% 
  select(-len_str)

```


#### English learners progress
```{r}
# No English learner variables for 2017.

# 2018
ELprog_18 <- read_excel("../data/source/report-cards/rc_2018.xlsx", sheet = 7) %>% 
  clean_names() %>% 
  select(1:5, "pct_met_prog")

# 2019
ELprog_19 <- read_excel("../data/source/report-cards/rc_2019.xlsx", sheet = 8) %>% 
  clean_names() %>% 
  select(1:5, "pct_met_prog")

# 2020
ELprog_20 <- read_excel("../data/source/report-cards/rc_2020.xlsx", sheet = 6) %>% 
  clean_names() %>%
  mutate(report_card_year = as.character(report_card_year), schoolid = as.character(schoolid), pct_met_prog = as.numeric(pct_met_prog)) %>% 
  select(1:5, "pct_met_prog") %>% 
  rename(schooltypecd = school_type_cd)

# 2021
ELprog_21 <- read_excel("../data/source/report-cards/rc_2021.xlsx", sheet = 9) %>% 
  clean_names() %>%
  mutate(report_card_year = as.character(report_card_year), schoolid = as.character(schoolid), pct_met_prog = as.numeric(pct_met_prog)) %>% 
  select(1:5, "pct_met_prog") %>% 
  rename(schooltypecd = school_type_cd)


# Bind English learners variables across years
ELprog_bind =  bind_rows(ELprog_18) %>% 
  bind_rows(ELprog_19) %>%
  bind_rows(ELprog_20) %>% 
  bind_rows(ELprog_21) %>%
  rename(any_of(key)) %>% 
  # Add leading zero to school ids with only 6 digits
  mutate(len_str = nchar(school_id)) %>%
  mutate(school_id = if_else(len_str == 6, paste0("0", school_id), school_id)) %>% 
  select(-len_str)
```

### Graduation rates
```{r}
grad_variables <- c("year", "district", "school", "bedscode", "gpercent_overall", "gpercent_race_w", "gpercent_race_b", "gpercent_race_h", "gpercent_race_i", "gpercent_race_a")


# 2017
grad_17 <- read_excel("../data/source/additional-info/grad-2017.xls") %>%
  clean_names() %>%
  mutate(year = "2017") %>% 
  select(any_of(grad_variables))

# 2018
grad_18 <- read_excel("../data/source/additional-info/grad-2018.xlsx") %>%
  clean_names() %>% 
  mutate(year = "2018") %>% 
  select(any_of(grad_variables))

# 2019
grad_19 <- read_excel("../data/source/additional-info/grad-2019.xlsx") %>%
  clean_names() %>%
  mutate(year = "2019") %>% 
  select(any_of(grad_variables))

# 2020
grad_20 <- read_excel("../data/source/additional-info/grad-2020.xlsx") %>%
  clean_names() %>% 
  mutate(year = "2020") %>% 
  select(any_of(grad_variables))

# 2021
grad_21 <- read_excel("../data/source/additional-info/grad-2021.xlsx") %>%
  clean_names() %>%
  mutate(year = "2021") %>% 
  select(any_of(grad_variables))


# Bind years
grad_bind = bind_rows(grad_17) %>% 
  bind_rows(grad_18) %>% 
  bind_rows(grad_19) %>% 
  bind_rows(grad_20) %>% 
  bind_rows(grad_21) %>%
  rename(any_of(key)) %>% 
  mutate(school_id = as.character(school_id), len_str = nchar(school_id)) %>% 
  mutate(school_id = if_else(len_str == 6, paste0("0", school_id), school_id)) %>% 
  select(-len_str) %>% 
  # Instead of a column identifying the institution type, these files have districts identified in the school name column as "DISTRICT WITH CHARTER" or "DISTRICT NO CHARTER." The first option matches the unique district ids in other files. Remove the second option.
  filter(school != "DISTRICT NO CHARTER")

```

#### Dropout rates
```{r}
# Define list of variables needed
dropout_vars <- c("report_card_year", "district_nm", "school_nm", "schoolid", "schooltypecd", "dropout_rate_curr_yr")

# 2017
dropout_dist_17 <- rc_dist_17 %>% 
  select(any_of(dropout_vars))

dropout_high_17 <- rc_high_17 %>% 
  select(any_of(dropout_vars))

# 2018
dropout_18 <- read_excel("../data/source/additional-info/ai-2018.xlsx", sheet = 4) %>%
  clean_names() %>% 
  select(any_of(dropout_vars))

# 2019 
dropout_19 <- read_excel("../data/source/additional-info/ai-2019.xlsx", sheet = 4) %>%
  clean_names() %>% 
  select(any_of(dropout_vars)) %>% 
  mutate(dropout_rate_curr_yr = as.numeric(dropout_rate_curr_yr))

# 2020 
dropout_20 <- read_excel("../data/source/additional-info/ai-2020.xlsx", sheet = 5) %>%
  clean_names() %>% 
  select(any_of(dropout_vars)) %>% 
  mutate(report_card_year = as.character(report_card_year), schoolid = as.character(schoolid))

# 2021 
dropout_21 <- read_excel("../data/source/additional-info/ai-2021.xlsx", sheet = 5) %>%
  clean_names() %>% 
  select(any_of(dropout_vars)) %>% 
  mutate(report_card_year = as.character(report_card_year), schoolid = as.character(schoolid))


# Bind years
dropout_bind = bind_rows(dropout_dist_17) %>% 
  bind_rows(dropout_high_17) %>% 
  bind_rows(dropout_18) %>% 
  bind_rows(dropout_19) %>% 
  bind_rows(dropout_20) %>% 
  bind_rows(dropout_21) %>% 
  rename(any_of(key)) %>% 
  mutate(len_str = nchar(school_id)) %>% 
  mutate(school_id = if_else(len_str == 6, paste0("0", school_id), school_id)) %>% 
  select(-len_str) %>% 
  distinct(.keep_all = TRUE)
```


### Load and clean climate data
#### Teachers
```{r}
# Define list of variables needed
teacher_variables <- c("report_card_year", "district_nm", "school_nm", "schoolid", "schooltypecd", "tchreturn3yr_avg_pct_curr_yr", "tchadvdegree_pct_curr_yr", "tchinexperienced_pct_curr_yr", "c23", "tchsalary_avg_curr_yr", "prime_instruct_time_curr_yr")

# 2017 - does not have comparable variables in teacher retention, advanced degrees or inexperienced teachers at the school level. It does have the percent of teachers with advanced degrees for districts and average teacher salary for schools and districts.

tch_dist_17 <- read_excel("../data/source/report-cards/rc_2017_district.xlsx", sheet = 2) %>%
  clean_names() %>% 
  mutate(report_card_year = "2017", schoolid = as.character(id), schooltypecd = type, tchreturn3yr_avg_pct_curr_yr = as.numeric(NA), tchinexperienced_pct_curr_yr = as.numeric(NA)) %>% 
  rename(district_nm = "distname", school_nm = "locname", tchadvdegree_pct_curr_yr = "c23", tchsalary_avg_curr_yr = "c29") %>% 
  select(any_of(teacher_variables))

tch_elem_17 <- read_excel("../data/source/report-cards/rc_2017_elem.xlsx", sheet = 2) %>%
  clean_names() %>% 
  mutate(report_card_year = "2017", schoolid = as.character(id), schooltypecd = type, tchreturn3yr_avg_pct_curr_yr = as.numeric(NA), tchinexperienced_pct_curr_yr = as.numeric(NA), tchadvdegree_pct_curr_yr = as.numeric(NA)) %>% 
  rename(district_nm = "distname", school_nm = "locname", tchsalary_avg_curr_yr = "c29") %>% 
  select(any_of(teacher_variables), -c23)

tch_middle_17 <- read_excel("../data/source/report-cards/rc_2017_middle.xlsx", sheet = 2) %>%
  clean_names() %>% 
  mutate(report_card_year = "2017", schoolid = as.character(id), schooltypecd = type, tchreturn3yr_avg_pct_curr_yr = as.numeric(NA), tchinexperienced_pct_curr_yr = as.numeric(NA), tchadvdegree_pct_curr_yr = as.numeric(NA)) %>% 
  rename(district_nm = "distname", school_nm = "locname", tchsalary_avg_curr_yr = "c29") %>% 
  select(any_of(teacher_variables), -c23)

tch_high_17 <- read_excel("../data/source/report-cards/rc_2017_high.xlsx", sheet = 2) %>%
  clean_names() %>% 
  mutate(report_card_year = "2017", schoolid = as.character(id), schooltypecd = type, tchreturn3yr_avg_pct_curr_yr = as.numeric(NA), tchinexperienced_pct_curr_yr = as.numeric(NA), tchadvdegree_pct_curr_yr = as.numeric(NA)) %>% 
  rename(district_nm = "distname", school_nm = "locname", tchsalary_avg_curr_yr = "c29") %>% 
  select(any_of(teacher_variables), -c23)

# 2018
tch_18 <- read_excel("../data/source/additional-info/ai-2018.xlsx", sheet = 9) %>%
  clean_names() %>% 
  select(any_of(teacher_variables))

# 2019
tch_19 <- read_excel("../data/source/additional-info/ai-2019.xlsx", sheet = 9) %>%
  clean_names() %>% 
  select(any_of(teacher_variables))

# 2020
tch_20 <- read_excel("../data/source/additional-info/ai-2020.xlsx", sheet = 10) %>%
  clean_names() %>% 
  select(any_of(teacher_variables)) %>% 
  mutate(report_card_year = as.character(report_card_year), tchreturn3yr_avg_pct_curr_yr = as.numeric(tchreturn3yr_avg_pct_curr_yr), tchinexperienced_pct_curr_yr = as.numeric(tchinexperienced_pct_curr_yr), tchsalary_avg_curr_yr = as.numeric(tchsalary_avg_curr_yr))

# 2021 
tch_21 <- read_excel("../data/source/additional-info/ai-2021.xlsx", sheet = 11) %>%
  clean_names() %>% 
  select(any_of(teacher_variables)) %>% 
  mutate(report_card_year = as.character(report_card_year), tchreturn3yr_avg_pct_curr_yr = as.numeric(tchreturn3yr_avg_pct_curr_yr), tchinexperienced_pct_curr_yr = as.numeric(tchinexperienced_pct_curr_yr), tchsalary_avg_curr_yr = as.numeric(tchsalary_avg_curr_yr))


# Bind years
teachers_bind = bind_rows(tch_dist_17) %>% 
  bind_rows(tch_elem_17) %>% 
  bind_rows(tch_middle_17) %>% 
  bind_rows(tch_high_17) %>% 
  bind_rows(tch_18) %>% 
  bind_rows(tch_19) %>% 
  bind_rows(tch_20) %>% 
  bind_rows(tch_21) %>% 
  rename(any_of(key)) %>% 
  mutate(len_str = nchar(school_id)) %>% 
  mutate(school_id = if_else(len_str == 6, paste0("0", school_id), school_id)) %>% 
  select(-len_str) %>% 
  distinct(.keep_all = TRUE)

```

#### Finances
```{r}
fin_variables <- c("report_card_year", "district_nm", "school_nm", "schoolid", "schooltypecd", "studentsin_poverty_pct_curr_yr", "expendfor_instruct_pct_curr_yr", "admin_salary_curr_yr", "total_ppe_curr_yr")

# 2017 - no admin salaries, per-pupil expenditures
fin_dist_17 <- read_excel("../data/source/report-cards/rc_2017_district.xlsx", sheet = 2) %>%
  clean_names() %>% 
  mutate(report_card_year = "2017", schoolid = as.character(id), schooltypecd = type) %>% 
  rename(district_nm = "distname", school_nm = "locname", studentsin_poverty_pct_curr_yr = "poverty", expendfor_instruct_pct_curr_yr = "c35") %>% 
  select(any_of(fin_variables))

fin_elem_17 <- read_excel("../data/source/report-cards/rc_2017_elem.xlsx", sheet = 2) %>%
  clean_names() %>% 
  mutate(report_card_year = "2017", schoolid = as.character(id), schooltypecd = type) %>% 
  rename(district_nm = "distname", school_nm = "locname", studentsin_poverty_pct_curr_yr = "poverty", expendfor_instruct_pct_curr_yr = "c35") %>% 
  select(any_of(fin_variables))

fin_middle_17 <- read_excel("../data/source/report-cards/rc_2017_middle.xlsx", sheet = 2) %>%
  clean_names() %>% 
  mutate(report_card_year = "2017", schoolid = as.character(id), schooltypecd = type) %>% 
  rename(district_nm = "distname", school_nm = "locname", studentsin_poverty_pct_curr_yr = "poverty", expendfor_instruct_pct_curr_yr = "c35") %>% 
  select(any_of(fin_variables))

fin_high_17 <- read_excel("../data/source/report-cards/rc_2017_high.xlsx", sheet = 2) %>%
  clean_names() %>% 
  mutate(report_card_year = "2017", schoolid = as.character(id), schooltypecd = type) %>% 
  rename(district_nm = "distname", school_nm = "locname", studentsin_poverty_pct_curr_yr = "poverty", expendfor_instruct_pct_curr_yr = "c35") %>% 
  select(any_of(fin_variables))


# 2018 - no per-pupil expenditures
fin_18 <- read_excel("../data/source/additional-info/ai-2018.xlsx", sheet = 11) %>%
  clean_names() %>% 
  select(any_of(fin_variables))

# 2019
fin_19 <- read_excel("../data/source/additional-info/ai-2019.xlsx", sheet = 11) %>%
  clean_names() %>% 
  select(any_of(fin_variables))

# 2020
fin_20 <- read_excel("../data/source/additional-info/ai-2020.xlsx", sheet = 12) %>%
  clean_names() %>% 
  select(any_of(fin_variables)) %>% 
  mutate(report_card_year = as.character(report_card_year), schoolid = as.character(schoolid), expendfor_instruct_pct_curr_yr = as.numeric(expendfor_instruct_pct_curr_yr), admin_salary_curr_yr = as.numeric(admin_salary_curr_yr), total_ppe_curr_yr = as.numeric(total_ppe_curr_yr))
         
# 2021 
fin_21 <- read_excel("../data/source/additional-info/ai-2021.xlsx", sheet = 13) %>%
  clean_names() %>% 
  select(any_of(fin_variables)) %>% 
  mutate(report_card_year = as.character(report_card_year), schoolid = as.character(schoolid), studentsin_poverty_pct_curr_yr = as.numeric(studentsin_poverty_pct_curr_yr), expendfor_instruct_pct_curr_yr = as.numeric(expendfor_instruct_pct_curr_yr), admin_salary_curr_yr = as.numeric(admin_salary_curr_yr), total_ppe_curr_yr = as.numeric(total_ppe_curr_yr))

# Bind years
finance_bind = bind_rows(fin_dist_17) %>% 
  bind_rows(fin_elem_17) %>% 
  bind_rows(fin_middle_17) %>% 
  bind_rows(fin_high_17) %>% 
  bind_rows(fin_18) %>% 
  bind_rows(fin_19) %>% 
  bind_rows(fin_20) %>% 
  bind_rows(fin_21) %>% 
  rename(any_of(key)) %>% 
  mutate(len_str = nchar(school_id)) %>% 
  mutate(school_id = if_else(len_str == 6, paste0("0", school_id), school_id)) %>% 
  select(-len_str) %>% 
  distinct(.keep_all = TRUE)

```

#### Technology
```{r}
tech_variables <- c("report_card_year", "district_nm", "school_nm", "schoolid", "schooltypecd", "gifted_pct_curr_yr", "pct78hs_credit_curr_yr", "princ_super_yrs_curr_yr", "tech_pct_class_wireless_curr_yr", "tech_oneto_one_curr_yr", "chronic_absenteeism_rate", "retained_pct_curr_yr")

# 2017
tech_dist_17 <- read_excel("../data/source/report-cards/rc_2017_district.xlsx", sheet = 2) %>%
  clean_names() %>% 
  mutate(report_card_year = "2017", schoolid = as.character(id), schooltypecd = type, cts_class = as.numeric(cts_class), cts_pctdev = as.numeric(cts_pctdev)) %>% 
  rename(district_nm = "distname", school_nm = "locname", tech_pct_class_wireless_curr_yr = "cts_class", tech_oneto_one_curr_yr = "cts_pctdev", retained_pct_curr_yr = "c3", gifted_pct_curr_yr = "c6", pct78hs_credit_curr_yr = "c1", princ_super_yrs_curr_yr = "c31") %>% 
  select(any_of(tech_variables))

tech_elem_17 <- read_excel("../data/source/report-cards/rc_2017_elem.xlsx", sheet = 2) %>%
  clean_names() %>% 
  mutate(report_card_year = "2017", schoolid = as.character(id), schooltypecd = type, cts_class = as.numeric(cts_class), cts_pctdev = as.numeric(cts_pctdev)) %>% 
  rename(district_nm = "distname", school_nm = "locname", tech_pct_class_wireless_curr_yr = "cts_class", tech_oneto_one_curr_yr = "cts_pctdev", retained_pct_curr_yr = "c3", gifted_pct_curr_yr = "c6", princ_super_yrs_curr_yr = "c31") %>% 
  select(any_of(tech_variables))

tech_middle_17 <- read_excel("../data/source/report-cards/rc_2017_middle.xlsx", sheet = 2) %>%
  clean_names() %>% 
  mutate(report_card_year = "2017", schoolid = as.character(id), schooltypecd = type, cts_class = as.numeric(cts_class), cts_pctdev = as.numeric(cts_pctdev)) %>% 
  rename(district_nm = "distname", school_nm = "locname", tech_pct_class_wireless_curr_yr = "cts_class", tech_oneto_one_curr_yr = "cts_pctdev", retained_pct_curr_yr = "c3", gifted_pct_curr_yr = "c6", pct78hs_credit_curr_yr = "c1", princ_super_yrs_curr_yr = "c31") %>% 
  select(any_of(tech_variables))

tech_high_17 <- read_excel("../data/source/report-cards/rc_2017_high.xlsx", sheet = 2) %>%
  clean_names() %>% 
  mutate(report_card_year = "2017", schoolid = as.character(id), schooltypecd = type, cts_class = as.numeric(cts_class), cts_pctdev = as.numeric(cts_pctdev)) %>% 
  rename(district_nm = "distname", school_nm = "locname", tech_pct_class_wireless_curr_yr = "cts_class", tech_oneto_one_curr_yr = "cts_pctdev", retained_pct_curr_yr = "c3", gifted_pct_curr_yr = "c6", pct78hs_credit_curr_yr = "c1", princ_super_yrs_curr_yr = "c31") %>% 
  select(any_of(tech_variables))

# chronic_absenteeism_rate = "" ?

# 2018
tech_18 <- read_excel("../data/source/additional-info/ai-2018.xlsx", sheet = 8) %>%
  clean_names() %>% 
  select(any_of(tech_variables))

# 2019
tech_19 <- read_excel("../data/source/additional-info/ai-2019.xlsx", sheet = 8) %>%
  clean_names() %>% 
  select(any_of(tech_variables))

# 2020
tech_20 <- read_excel("../data/source/additional-info/ai-2020.xlsx", sheet = 9) %>%
  clean_names() %>% 
  select(any_of(tech_variables)) %>% 
  mutate(report_card_year = as.character(report_card_year), schoolid = as.character(schoolid), gifted_pct_curr_yr = as.numeric(gifted_pct_curr_yr), pct78hs_credit_curr_yr = as.numeric(pct78hs_credit_curr_yr), princ_super_yrs_curr_yr = as.numeric(princ_super_yrs_curr_yr), tech_pct_class_wireless_curr_yr = as.numeric(tech_pct_class_wireless_curr_yr), tech_oneto_one_curr_yr = as.numeric(tech_oneto_one_curr_yr), chronic_absenteeism_rate = as.numeric(chronic_absenteeism_rate))

# 2021
tech_21 <- read_excel("../data/source/additional-info/ai-2021.xlsx", sheet = 9) %>%
  clean_names() %>% 
  select(any_of(tech_variables)) %>%
  mutate(report_card_year = as.character(report_card_year), schoolid = as.character(schoolid), gifted_pct_curr_yr = as.numeric(gifted_pct_curr_yr), pct78hs_credit_curr_yr = as.numeric(pct78hs_credit_curr_yr), princ_super_yrs_curr_yr = as.numeric(princ_super_yrs_curr_yr), tech_pct_class_wireless_curr_yr = as.numeric(tech_pct_class_wireless_curr_yr), tech_oneto_one_curr_yr = as.numeric(tech_oneto_one_curr_yr), chronic_absenteeism_rate = as.numeric(chronic_absenteeism_rate), retained_pct_curr_yr = as.numeric(retained_pct_curr_yr))


# Bind years
tech_bind = tech_dist_17 %>%
  bind_rows(tech_elem_17) %>%
  bind_rows(tech_middle_17) %>%
  bind_rows(tech_high_17) %>% 
  bind_rows(tech_18) %>% 
  bind_rows(tech_19) %>% 
  bind_rows(tech_20) %>% 
  bind_rows(tech_21) %>%
  rename(any_of(key)) %>%
  mutate(len_str = nchar(school_id)) %>% 
  mutate(school_id = if_else(len_str == 6, paste0("0", school_id), school_id)) %>% 
  select(-len_str) %>%
  # Some schools are duplicated on all variables. Remove duplicates.
  distinct(.keep_all = TRUE)

```


#### Chronic absenteeism, English learners, students with disabilities
```{r}
stu_vars <- c("report_card_year", "district_nm", "school_nm", "schoolid", "schooltypecd", "chronic_absenteeism_rate", "n_total_di", "n_total_el")

# 2017 - not available (Variables location sheet says C4 is chronic absenteeism rate but the actual spreadsheet says C4 is "Attendance rate - our school.")

# 2018
stu_18 <- read_excel("../data/source/additional-info/ai-2018.xlsx", sheet = 8) %>%
  clean_names() %>% 
  select(any_of(stu_vars))

# 2019 - not available

# 2020 - not available

# 2021
stu_21 <- read_excel("../data/source/additional-info/ai-2021.xlsx", sheet = 10) %>%
  clean_names() %>% 
  rename(chronic_absenteeism_rate = "pct_chronic_all") %>% 
  select(any_of(stu_vars)) %>% 
  mutate(report_card_year = as.character(report_card_year), schoolid = as.character(schoolid), n_total_el = as.numeric(n_total_el))

  
# Bind years
students_bind = bind_rows(stu_18) %>%
  bind_rows(stu_21) %>%
  rename(any_of(key)) %>%
  mutate(len_str = nchar(school_id)) %>% 
  mutate(school_id = if_else(len_str == 6, paste0("0", school_id), school_id)) %>% 
  select(-len_str) %>%
  # Some schools are duplicated on all variables. Remove duplicates.
  distinct(.keep_all = TRUE)

```


#### Safety
```{r}
safety_vars <- c("report_card_year", "district_nm", "school_nm", "schoolid", "schooltypecd", "par_child_feels_safe", "par_tch_stop_bully", "par_num_responses", "tch_i_feel_safe", "tch_rules_enforced", "tch_num_responses", "ocr_firearm_num", "ocr_rape", "ocr_sex_assault", "ocr_robw_weapon", "ocr_robw_firearm", "ocr_rob_no_weapon", "ocr_phys_attachw_weapon", "ocr_phys_attackw_firearm", "ocr_phys_attach_no_weapon", "ocr_threat_phys_attachw_weapon", "ocr_threat_phys_attackw_firearm", "ocr_threat_phys_attach_no_weapon", "ocr_in_school_suspension", "ocr_out_of_school_suspension", "ocr_expelled", "ocr_arrests_referrals", "ocr_bully_harass")


# 2017 - not available

# 2018 - Suspensions, expulsions, bullying incidents, school related arrests and referrals to law enforcement not available
safety_18 <- read_excel("../data/source/additional-info/ai-2018.xlsx", sheet = 10) %>%
  clean_names() %>% 
  select(any_of(safety_vars)) %>% 
  mutate(par_child_feels_safe = as.numeric(par_child_feels_safe), par_tch_stop_bully = as.numeric(par_tch_stop_bully), par_num_responses = as.numeric(par_num_responses), tch_i_feel_safe = as.numeric(tch_i_feel_safe), tch_rules_enforced = as.numeric(tch_rules_enforced), tch_num_responses = as.numeric(tch_rules_enforced))

# 2019
safety_19 <- read_excel("../data/source/additional-info/ai-2019.xlsx", sheet = 10) %>%
  clean_names() %>% 
  select(any_of(safety_vars))

# 2020 - parent and teacher opinions not available
safety_20 <- read_excel("../data/source/additional-info/ai-2020.xlsx", sheet = 11) %>%
  clean_names() %>%
  rename(schoolid = "sidn") %>% 
  select(any_of(safety_vars)) %>% 
  mutate(report_card_year = as.character(report_card_year), ocr_firearm_num = as.numeric(ocr_firearm_num), ocr_rape = as.numeric(ocr_rape), ocr_sex_assault = as.numeric(ocr_sex_assault), ocr_robw_weapon = as.numeric(ocr_robw_weapon), ocr_robw_firearm = as.numeric(ocr_robw_firearm), ocr_rob_no_weapon = as.numeric(ocr_rob_no_weapon), ocr_phys_attachw_weapon = as.numeric(ocr_phys_attachw_weapon), ocr_phys_attackw_firearm = as.numeric(ocr_phys_attackw_firearm), ocr_phys_attach_no_weapon = as.numeric(ocr_phys_attach_no_weapon), ocr_threat_phys_attachw_weapon = as.numeric(ocr_threat_phys_attachw_weapon), ocr_threat_phys_attackw_firearm = as.numeric(ocr_threat_phys_attackw_firearm), ocr_threat_phys_attach_no_weapon = as.numeric(ocr_threat_phys_attach_no_weapon), ocr_in_school_suspension = as.numeric(ocr_in_school_suspension), ocr_out_of_school_suspension = as.numeric(ocr_out_of_school_suspension), ocr_expelled = as.numeric(ocr_expelled), ocr_arrests_referrals = as.numeric(ocr_arrests_referrals), ocr_bully_harass = as.numeric(ocr_bully_harass))


# 2021 
safety_21 <- read_excel("../data/source/additional-info/ai-2021.xlsx", sheet = 12) %>%
  clean_names() %>% 
  rename(schoolid = "sidn") %>% 
  select(any_of(safety_vars)) %>% 
  mutate(report_card_year = as.character(report_card_year), par_child_feels_safe = as.numeric(par_child_feels_safe), par_tch_stop_bully = as.numeric(par_tch_stop_bully), par_num_responses = as.numeric(par_num_responses), tch_i_feel_safe = as.numeric(tch_i_feel_safe), tch_rules_enforced = as.numeric(tch_rules_enforced), tch_num_responses = as.numeric(tch_rules_enforced), ocr_firearm_num = as.numeric(ocr_firearm_num), ocr_rape = as.numeric(ocr_rape), ocr_sex_assault = as.numeric(ocr_sex_assault), ocr_robw_weapon = as.numeric(ocr_robw_weapon), ocr_robw_firearm = as.numeric(ocr_robw_firearm), ocr_rob_no_weapon = as.numeric(ocr_rob_no_weapon), ocr_phys_attachw_weapon = as.numeric(ocr_phys_attachw_weapon), ocr_phys_attackw_firearm = as.numeric(ocr_phys_attackw_firearm), ocr_phys_attach_no_weapon = as.numeric(ocr_phys_attach_no_weapon), ocr_threat_phys_attachw_weapon = as.numeric(ocr_threat_phys_attachw_weapon), ocr_threat_phys_attackw_firearm = as.numeric(ocr_threat_phys_attackw_firearm), ocr_threat_phys_attach_no_weapon = as.numeric(ocr_threat_phys_attach_no_weapon), ocr_in_school_suspension = as.numeric(ocr_in_school_suspension), ocr_out_of_school_suspension = as.numeric(ocr_out_of_school_suspension), ocr_expelled = as.numeric(ocr_expelled), ocr_arrests_referrals = as.numeric(ocr_arrests_referrals), ocr_bully_harass = as.numeric(ocr_bully_harass))


# Bind years
safety_bind = bind_rows(safety_18) %>% 
  bind_rows(safety_19) %>% 
  bind_rows(safety_20) %>% 
  bind_rows(safety_21) %>% 
  rename(any_of(key)) %>% 
  # All school ids are 6 characters. Do not need to introduce leading 0s. Also, no duplicates, so do not need to distinct.
  # mutate(len_str = nchar(school_id)) %>% 
  # mutate(school_id = if_else(len_str == 6, paste0("0", school_id), school_id)) %>% 
  # select(-len_str) %>%
  # distinct(.keep_all = TRUE) %>% 
  # Sum all violent incidents
  mutate(ocr_violent_all = ocr_rape + ocr_sex_assault + ocr_robw_weapon + ocr_robw_firearm + ocr_phys_attachw_weapon + ocr_phys_attackw_firearm + ocr_threat_phys_attachw_weapon + ocr_threat_phys_attackw_firearm) %>% 
  mutate(ocr_nonviolent_all = ocr_rob_no_weapon + ocr_phys_attach_no_weapon + ocr_threat_phys_attach_no_weapon)
  
```


### Load and clean enrollment and demographic data
#### Schools
```{r}
# 2017
school_hc_17 <- read_excel("../data/source/headcounts/2016_17_school_headcount_demographics.xlsx", col_names = c("school_id", "district", "school", "total_actively_enrolled_students", "gender_female", "gender_male", "black_or_african_american", "american_indian", "asian_hawaiian_pacific_islander", "hispanic", "white", "race_missing"), sheet = 1) %>%
  slice(7:1218) %>%
  mutate(year = "2017")


# 2018
school_hc_18 <- read_excel("../data/source/headcounts/2017_18_school_headcount_demographics.xlsx", col_names = c("school_id", "district", "school", "total_actively_enrolled_students", "empty1", "gender_female", "gender_male", "gender_missing", "empty2", "black_or_african_american", "american_indian", "asian", "hispanic_or_latino", "hawaiian_or_other_pacific_islander", "two_or_more_races", "white", "race_missing"), sheet = 1) %>%
  remove_empty(c('rows', "cols")) %>% 
  select(-empty2) %>% 
  slice(8:1220) %>%
  mutate(year = "2018")


# 2019
school_hc_19 <- read_excel("../data/source/headcounts/2018_19_school_headcount_demographics.xlsx", col_names = c("school_id", "district", "school", "total_actively_enrolled_students", "gender_female", "gender_male", "gender_missing", "black_or_african_american", "american_indian", "asian", "hispanic_or_latino", "hawaiian_or_other_pacific_islander", "two_or_more_races", "white", "race_missing", "pupils_in_poverty"), sheet = 1) %>%
  slice(8:1227) %>% 
  select(-pupils_in_poverty) %>% 
  mutate(year = "2019")


# 2020
school_hc_20 <- read_excel("../data/source/headcounts/2019_20_school_headcount_demographics.xlsx", col_names = c("school_id", "district", "school", "total_actively_enrolled_students", "gender_female", "gender_male", "gender_missing", "black_or_african_american", "american_indian", "asian", "hispanic_or_latino", "hawaiian_or_other_pacific_islander", "two_or_more_races", "white", "race_missing", "pupils_in_poverty"), sheet = 1) %>%
  slice(8:1229) %>% 
  select(-pupils_in_poverty) %>% 
  mutate(year = "2020")


# 2021 loaded above (for use in creating school list)


# Bind across years for schools
school_bind = school_hc_17 %>% 
  bind_rows(school_hc_18) %>% 
  bind_rows(school_hc_19) %>% 
  bind_rows(school_hc_20) %>% 
  bind_rows(school_hc_21) %>% 
  mutate(total_actively_enrolled_students = as.integer(total_actively_enrolled_students),
         gender_female = as.integer(gender_female),
         gender_male = as.integer(gender_male),
         gender_missing = as.integer(gender_missing),
         black_or_african_american = as.integer(black_or_african_american),
         american_indian = as.integer(american_indian),
         asian = as.integer(asian),
         hispanic_or_latino = as.integer(hispanic_or_latino),
         hawaiian_or_other_pacific_islander = as.integer(hawaiian_or_other_pacific_islander),
         two_or_more_races = as.integer(two_or_more_races),
         white = as.integer(white),
         race_missing = as.integer(race_missing),
         asian_hawaiian_pacific_islander = as.integer(asian_hawaiian_pacific_islander),
         hispanic = as.integer(hispanic)
         )

# Make final headcount df
school_hc_all_years <- school_bind %>% 
  # Sum across certain races because racial categories set by the state differ across years. P&C team opted for combining races other than white and Black into one "other races" category. Also, in this step we will lose data on students whose race was categorized as "missing," so the sum across ALL races may not equal the total_actively_enrolled students. Across the entire state, the number of students in this category was no higher than XX in any of the years we are examining.
  mutate(all_other_races = rowSums(school_bind[,c("american_indian","asian", "hispanic_or_latino","hawaiian_or_other_pacific_islander","two_or_more_races","asian_hawaiian_pacific_islander","hispanic")], na.rm=TRUE)) %>% 
  # Deselect unneeded columns
  select(-american_indian,-asian,-hispanic_or_latino,-hawaiian_or_other_pacific_islander,-two_or_more_races,-race_missing,-asian_hawaiian_pacific_islander,-hispanic)  %>% 
  # Calculate percentages
  mutate(percent_female = gender_female/total_actively_enrolled_students*100, percent_male = gender_male/total_actively_enrolled_students*100, percent_gender_missing = gender_missing/total_actively_enrolled_students*100, percent_black_aa = black_or_african_american/total_actively_enrolled_students*100, percent_white = white/total_actively_enrolled_students*100, percent_all_other_races = all_other_races/total_actively_enrolled_students*100)

```


### District enrollment and demographics
```{r}
# 2017
district_hc_17 <- read_excel("../data/source/headcounts/2016_17_district_headcount_demographics.xlsx", col_names = c("extra", "district_id", "district", "total_actively_enrolled_students", "gender_female", "gender_male", "black_or_african_american", "american_indian", "asian_or_pacific_islander", "hispanic", "white", "race_missing"), sheet = 1) %>%
  select(-"extra") %>% 
  slice(7:93) %>%
  mutate(year = "2017")


# 2018
district_hc_18 <- read_excel("../data/source/headcounts/2017_18_district_headcount_demographics.xlsx", col_names = c("district_id", "district", "total_actively_enrolled_students", "empty1", "gender_female", "gender_male", "gender_missing", "empty2", "black_or_african_american", "american_indian", "asian", "hispanic_or_latino", "hawaiian_or_other_pacific_islander", "two_or_more_races", "white", "race_missing"), sheet = 1) %>%
  remove_empty(c('rows', "cols")) %>% 
  slice(8:94) %>%
  mutate(year = "2018") %>% 
  mutate(schoolid = paste0(district_id, "999"))
  

# 2019
district_hc_19 <- read_excel("../data/source/headcounts/2018_19_district_headcount_demographics.xlsx", col_names = c("district_id", "district", "total_actively_enrolled_students", "gender_female", "gender_male", "gender_missing", "black_or_african_american", "american_indian", "asian", "hispanic_or_latino", "hawaiian_or_other_pacific_islander", "two_or_more_races", "white", "race_missing", "pupils_in_poverty"), sheet = 1) %>%
  slice(8:95) %>%
  select(-pupils_in_poverty) %>% 
  mutate(year = "2019")


# 2020
district_hc_20 <- read_excel("../data/source/headcounts/2019_20_district_headcount_demographics.xlsx", col_names = c("district_id", "district", "total_actively_enrolled_students", "gender_female", "gender_male", "gender_missing", "black_or_african_american", "american_indian", "asian", "hispanic_or_latino", "hawaiian_or_other_pacific_islander", "two_or_more_races", "white", "race_missing", "pupils_in_poverty"), sheet = 1) %>%
  slice(8:93) %>%
  select(-pupils_in_poverty) %>% 
  mutate(year = "2020")


# 2021
district_hc_21 <- read_excel("../data/source/headcounts/2020_21_district_headcount_demographics.xlsx", col_names = c("district_id", "district", "total_actively_enrolled_students", "gender_female", "gender_male", "gender_missing", "black_or_african_american", "american_indian", "asian", "hispanic_or_latino", "hawaiian_or_other_pacific_islander", "two_or_more_races", "white", "race_missing", "pupils_in_poverty"), sheet = 1) %>%
  slice(8:94) %>% 
  select(-pupils_in_poverty) %>% 
  mutate(year = "2021")


#### Create a dataframe with district headcounts for all years
# Bind districts across years
district_bind = district_hc_17 %>%
  bind_rows(district_hc_18) %>%
  bind_rows(district_hc_19) %>% 
  bind_rows(district_hc_20) %>% 
  bind_rows(district_hc_21) %>% 
  # Add trailing "999" to match ids in other datasets
  mutate(district_id = paste0(district_id, "999")) %>%
  # Remove unneeded districts
  filter(!(district_id %in% drop_districts)) %>% 
  # Many of the columns are stored as character values. Convert to numbers.
  mutate(total_actively_enrolled_students = as.numeric(total_actively_enrolled_students),
         gender_female = as.numeric(gender_female),
         gender_male = as.numeric(gender_male),
         gender_missing = as.numeric(gender_missing),
         black_or_african_american = as.numeric(black_or_african_american),
         american_indian = as.numeric(american_indian),
         asian = as.numeric(asian),
         hispanic_or_latino = as.numeric(hispanic_or_latino),
         hawaiian_or_other_pacific_islander = as.numeric(hawaiian_or_other_pacific_islander),
         two_or_more_races = as.numeric(two_or_more_races),
         white = as.numeric(white),
         race_missing = as.numeric(race_missing),
         asian_or_pacific_islander = as.numeric(asian_or_pacific_islander),
         hispanic = as.numeric(hispanic)
         )


# Make final df
district_hc_all_years <- district_bind %>%
  # Sum across certain races because racial categories set by the state differ across years. P&C team opted for combining races other than white and Black into one "other races" category. Also, in this step we will lose data on students whose race was categorized as "missing," so the sum across ALL races may not equal the total_actively_enrolled students. Across the entire state, the number of students in this category was no higher than XX in any of the years we are examining.
  mutate(all_other_races = rowSums(district_bind[,c("american_indian","asian", "hispanic_or_latino","hawaiian_or_other_pacific_islander","two_or_more_races", "asian_or_pacific_islander","hispanic")], na.rm=TRUE)) %>%
  # Deselect unneeded columns
  select(-american_indian,-asian,-hispanic_or_latino,-hawaiian_or_other_pacific_islander,-two_or_more_races,-race_missing,-asian_or_pacific_islander,-hispanic) %>% 
  # Calculate percentages
  mutate(percent_female = gender_female/total_actively_enrolled_students*100, percent_male = gender_male/total_actively_enrolled_students*100, percent_gender_missing = gender_missing/total_actively_enrolled_students*100, percent_black_aa = black_or_african_american/total_actively_enrolled_students*100, percent_white = white/total_actively_enrolled_students*100, percent_all_other_races = all_other_races/total_actively_enrolled_students*100)


```


### Join all variables
```{r}
# Join academic variables
academic_variables <- acad_em_bind %>%
  # Names match in these two dfs, so no need to specify which columns we're joining on
  full_join(acad_high_bind) %>% 
  # Other variables come from different spreadsheets. Specify join columns
  full_join(act_bind, by = c("year", "school_id", "type")) %>% 
  full_join(career_bind, by = c("year", "school_id", "type")) %>% 
  full_join(ELprog_bind, by = c("year", "school_id", "type")) %>% 
  full_join(grad_bind, by = c("year", "school_id")) %>% 
  full_join(dropout_bind, by = c("year", "school_id", "type")) %>% 
  # Remove school and district names in prep for joining with other variable dfs
  select(-school.x.x.x, -district.x.x.x, -school.x, -district.x, -school.y, -district.y, -school.x.x, -district.x.x, -school.y.y, -district.y.y)


# Checks - grad, act, career, ELprog_bind (primaries, districts 2020), grad (districts 2017)
# anti <- grad_bind %>% 
#   anti_join(academic_variables, by = c("year", "school_id"))

# Join climate variables
climate_variables <- teachers_bind %>% 
  full_join(tech_bind, by = c("year", "school_id", "type")) %>% 
  full_join(finance_bind, by = c("year", "school_id", "type")) %>% 
  full_join(students_bind, by = c("year", "school_id", "type")) %>%
  full_join(safety_bind, by = c("year", "school_id", "type")) %>% 
  # Remove school and district names in prep for joining with other variable dfs
  select(-school.x, -district.x, -school.y, -district.y, -school.x.x, -district.x.x, -school.y.y, -district.y.y)
  

# Checks - 
# anti <- teachers_bind %>%
#   anti_join(tech_bind, by = c("year", "school_id", "type"))

# Join academic df, climate df
all_variables <- academic_variables %>% 
  full_join(climate_variables, by = c("school_id", "year", "type"))


## Create separate dataframes for districts and schools and join with headcounts
# Districts 
district_av <- all_variables %>%
  filter(type == "D") %>% 
  rename(district_id = school_id) %>% 
  # Join with headcount
  left_join(district_hc_all_years, by = c("district_id", "year"))

  ### Some with missing years?

# Schools
school_av <- all_variables %>%
  filter(type != "D", type !="S") %>% 
  # Join with headcount -- challenge in that there's no "type" column ----------------- Maybe school and district lists should be joined with headcounts before joining with other variables??
  left_join(school_hc_all_years, by = c("school_id", "year"))


```




# Join with canonical lists. This will remove unnecessary schools and bring in address data.
```{r}
# District
district_full_data <- district_list %>% 
  left_join(district_av, by = "district_id") %>% 
  # Calculate percentages of English language learners and students without (the latter is necessary for charts)
  mutate(pct_el = (n_total_el/enrollment)*100, pct_non_el = 100-pct_el) %>%  
  # Calculate percentages of students with disabilities and students without (the latter is necessary for charts)
  mutate(pct_disabilities = (n_total_di/enrollment)*100, pct_non_di = 100-pct_disabilities) %>%
  # Calculate percentages of students not in poverty (necessary for charts)
  mutate(pct_not_in_poverty = 100-studentsin_poverty_pct_curr_yr) %>% 
  # Remove carriage returns
  clean_names() %>% 
  select(-district_y_y_y,-school_y_y_y,-district_y,-district_x,-school_x) %>% 
  mutate(district = gsub("\r?\n|\r", " ", district)) %>% 
  mutate(district_alt = gsub("\r?\n|\r", " ", district_alt)) %>% 
  mutate(supername = gsub("\r?\n|\r", " ", supername)) %>% 
  mutate(street = gsub("\r?\n|\r", " ", street)) %>% 
  mutate(city = gsub("\r?\n|\r", " ", city)) %>% 
  mutate(chair = gsub("\r?\n|\r", " ", chair)) %>% 
  select(year, 1:14,16:118) %>% 
  select(-school_y)
write_csv(district_full_data, "../data/processed/etl_district_full_data.csv")

# School
school_full_data <- school_list_geocoded %>% 
  left_join(school_av, by = "school_id") %>% 
  # Calculate percentages of English language learners and students without (the latter is necessary for charts)
  mutate(pct_el = (n_total_el/enrollment)*100, pct_non_el = 100-pct_el) %>%  
  # Calculate percentages of students with disabilities and students without (the latter is necessary for charts)
  mutate(pct_disabilities = (n_total_di/enrollment)*100, pct_non_di = 100-pct_disabilities) %>%
  # Calculate percentages of students not in poverty (necessary for charts)
  mutate(pct_not_in_poverty = 100-studentsin_poverty_pct_curr_yr) %>% 
  # Remove carriage returns
  clean_names() %>%
  select(-district_y_y_y,-school_y_y_y,-district_y,-district_x,-school_x,-school_y) %>% 
  mutate(school = gsub("\r?\n|\r", " ", school)) %>% 
  mutate(district = gsub("\r?\n|\r", " ", district)) %>% 
  mutate(district_alt = gsub("\r?\n|\r", " ", district_alt)) %>% 
  mutate(street = gsub("\r?\n|\r", " ", street)) %>% 
  mutate(hdphone = gsub("\r?\n|\r", " ", hdphone)) %>% 
  mutate(url = gsub("\r?\n|\r", " ", url)) %>% 
  mutate(city = gsub("\r?\n|\r", " ", city)) 

  
```



## Make key to variable names
```{r}
vars_d = colnames(district_full_data)
variables_district = data.frame(variables = vars_d)

vars_s = colnames(school_full_data)
variables_school = data.frame(variables = vars_s)

```



## Export data
```{r}
# Canonical lists
write_csv(district_list, "../data/processed/etl_district_list.csv")
write_csv(school_list, "../data/processed/etl_school_list.csv")

# Full data
write_csv(district_full_data, "../data/processed/etl_district_full_data.csv")
write_csv(school_full_data, "../data/processed/etl_school_full_data.csv")

# Variable names
write_csv(variables_district, "../data/processed/etl_variables_district.csv")
write_csv(variables_school, "../data/processed/etl_variables_school.csv")



```

-30-